---
title: '02.3 Practice'
author: "R13B42015"
date: "2024-10-05"
output: html_document
---

```{r eval = TRUE,  message=F, include=FALSE, warning=F, purl=F, results="hide"}
#knitr::opts_chunk$set(echo = TRUE)
```

## Download rairuoho dataset
Click the button below to download the dataset. It contains data on 8 days length of Easter grass exposed to different treatments. 
```{r echo = FALSE }
xfun::embed_file("C:\\ocean5098\\02.3-Rairuoho\\rairuoho_data.txt")
```

```{r, message=FALSE, warning=FALSE}

```

```{r ,eval=FALSE}
library(dplyr)
library(tidyr)
```

## Read rairuoho dataset
```{r}
rairuoho_path<- "C:\\ocean5098\\02.3-Rairuoho\\rairuoho_data.txt"
rairuoho_data<- read.table(rairuoho_path, header = T)
head(rairuoho_data,5)
```

## Replace nutrient with enriched in the dataset
```{r}
rairuoho_data[rairuoho_data$treatment == "nutrient","treatment"]<- "enriched"
head(rairuoho_data,5)
```
## Reformat dataset
```{r}
rairuoho_re <- reshape(rairuoho_data,
                         varying = list(names(rairuoho_data)[grepl("day", names(rairuoho_data))]),
                         v.names = "length",
                         timevar = "day",
                         times = c("day3", "day4", "day5", "day6", "day7", "day8"),
                         direction = "long")
rairuoho_re <- rairuoho_re[ , !names(rairuoho_re) %in% c("id")]
rownames(rairuoho_re)<-NULL
head(rairuoho_re,5)
```

## set factor
let column "day" in dataset be a factor with 6 levels
```{r}
rairuoho_re$day <- factor(rairuoho_re$day, levels = c("day3", "day4", "day5", "day6", "day7", "day8"))
head(rairuoho_re,5)
str(rairuoho_re)
```

## merge spatial
merge two spatial with "_" and replace two column called "spatial1" and "spatial1", and delete two column called "row" and "column" in dataset
```{r}
head(rairuoho_re$spatial1)    # spatial 1
head(rairuoho_re$spatial2)    # spatial 2
spatial<- paste(rairuoho_re$spatial1, rairuoho_re$spatial2, sep = "_")
rairuoho_spat<- cbind(rairuoho_re,spatial)
head(rairuoho_spat,5)
```

```{r}
rairuoho <- rairuoho_spat[ , !names(rairuoho_spat) %in% c("spatial1", "spatial2", "row", "column")]
head(rairuoho,5)
```

## save as csv file
```{r}
csv_path<- "C:\\ocean5098\\02.3-Rairuoho\\rairuoho.csv"
write.csv(rairuoho, csv_path, row.names=F)
```
```{r echo = FALSE }
xfun::embed_file(csv_path)
```