---
title: 'practice 3.2'
author: "R13B42015"
date: "2024-10-13"
output: html_document
---

```{r ,eval = TRUE,  message=F, include=FALSE, warning=F, purl=F, results="hide"}
#knitr::opts_chunk$set(echo = TRUE)
```
# the first element of the "my.experiment" list
The data frame contains the growth conditions of five individuals.
```{r}
start_experiment <-  c(4, 5, 3, 6, 7)
end_experiment <-  c(9, 6, 5, 4, 7)
plant_data <- data.frame(start_experiment, end_experiment)
row.names(plant_data) <-  c("plant 1", "plant 2", "plant 3", "plant 4", "plant 5")
plant_data
```

# Reformat the data frame named "plant_data"
```{r}
library(tidyr)
long_plant_data <- pivot_longer(plant_data,
                                cols = c(start_experiment,
                                         end_experiment),
                                names_to = "time",
                                values_to = "height")
long_plant_data
str(long_plant_data)
long_plant_data$time <- factor(long_plant_data$time, levels = c("start_experiment", "end_experiment"))
long_plant_data$time
str(long_plant_data$time)
```

# the second element of my.experiment list
### the first element of growth list
Extract 5 row names from "plant_data".
```{r}
row_names <- rownames(plant_data)
```
### the second element of growth list
Numeric vector of percentage growth.
```{r}
initial_heights <- plant_data$start_experiment
initial_heights
final_heights <- plant_data$end_experiment
final_heights
percentage_growth <- ((final_heights - initial_heights) / initial_heights) * 100
percentage_growth
```
### the third element of growth list
Data frame containing plants and their growth
```{r}
growth_table <- data.frame(plants = row_names,
                           growth = percentage_growth)
growth_table
```
### Create the growth list.
```{r}
growth_list <- list(row_names, percentage_growth, growth_table)
growth_list
```

# my.experiment list
Combine the three elements—long_plant_data, growth_list, and a message about the growth status—into a single list titled my.experiment.
```{r}
my.experiment <- list(long_plant_data, 
                      growth_list, 
                      "The growth of the plants varies; some increased, while others decreased.")
my.experiment
```
