---
title: "Practice 7.4 - LV"
author: "R13B42015"
date: "2024-11-17"
output: html_document
---

```{r ,eval = TRUE,  message=F, include=FALSE, warning=F, purl=F, results="hide"}
#knitr::opts_chunk$set(echo = TRUE)
```

```{r ,message=F}
library(dplyr)
library(tidyr)
library(ggplot2)
```

# Define the correlation function
```{r}
cor_func<- function(dataset, day_start, day_end, treatment)
{
  #############################
  # set vector of test data
  treatment_start<- dataset[dataset$treatment == treatment, day_start]
  treatment_end<- dataset[dataset$treatment == treatment, day_end]
  
  #############################
  # correlation (r value)
  r<- cor(treatment_start, treatment_end)
  #############################
  # t value
  n <- length(treatment_start)  # 假設樣本數
  t_value <- r * sqrt(n - 2) / sqrt(1 - r^2)
  #############################
  # p value
  degree_of_freedom<- n - 2
  p_value <- 2 * (1 - pt(abs(t_value), df = degree_of_freedom))
  print(p_value)
  
  cat("data: gress length of ", day_start," and ",day_end, "\n",
      "treatment: ", treatment, "\n",
      "t = ", t_value, "\n",
      "p value = ", p_value, "\n",
      "correlation: ", r)
  
  return(data.frame(day_start = day_start, day_end = day_end, 
                    treatment = treatment, correlation = r, 
                    t = t_value, p_value = p_value))
}
```

# read dataset
```{r}
rairuoho <- read.table("C:\\ocean5098\\02.3-Rairuoho\\rairuoho_data.txt", header = TRUE)
head(rairuoho,5)
```

# Create Boxplots for Grass Length Comparison
```{r}
rairuoho_long <- pivot_longer(rairuoho, cols = c(day3, day4), names_to = "day", values_to = "length")
ggplot(rairuoho_long, aes(x = day, y = length, fill = treatment)) +
  geom_boxplot() +
  labs(title = "Comparison of grass length on day 3 and day 8 between different treatments",
       x = "day",
       y = "length")

rairuoho_long <- pivot_longer(rairuoho, cols = c(day3, day8), names_to = "day", values_to = "length")
ggplot(rairuoho_long, aes(x = day, y = length, fill = treatment)) +
  geom_boxplot() +
  labs(title = "Comparison of grass length on day 3 and day 8 between different treatments",
       x = "day",
       y = "length")
```

# Perform Correlation Tests
```{r}
treatments <- c("water", "nutrient")
day_pairs <- list(c("day3", "day4"), c("day3", "day8"))
results <- data.frame()
for(treatment in treatments)
{
  for(day_pair in day_pairs)
  {
    result<- cor_func(rairuoho, day_pair[1], day_pair[2], treatment)
    results <- rbind(results, result)
  }
}
results
```

# Results:
1. Water Treatment<br>
  **Correlation between Day 3 and Day 4**<br>
    Correlation coefficient ( r = 0.7865 ): This indicates a strong positive correlation between the lengths on Day 3 and Day 4.
    p value = 5.197822e-06,  this value is less than 0.01, showing that this correlation is statistically very significant.<br>
  **Correlation between Day 3 and Day 8**<br>
    Correlation coefficient ( r = 0.4596 ): This shows a moderate positive correlation between lengths on Day 3 and Day 8.
    p value = 2.386844e-02, this value is less than 0.05, indicating that this correlation is also significant, but the significance is weaker compared to the first pair of comparisons.

2. Nutrient Treatment<br>
  **Correlation between Day 3 and Day 4**<br>
    Correlation coefficient ( r = 0.7744 ): This shows a strong positive correlation, similar to the water treatment group.
    p value = 8.938370e-06, this value also shows high significance.<br>
  **Correlation between Day 3 and Day 8**<br>
    	Correlation coefficient ( r = 0.2417 ): This value indicates a weak positive correlation.p value = 2.552119e-01, this value is greater than 0.05, suggesting that this correlation is not statistically significant.


# Conclusion
From the above analysis, we observe that:<br>
<div style="border: 2px solid #0073e6; padding: 10px; border-radius: 5px; background-color: #f9f9f9;">
•	The Water Treatment shows a strong and significant correlation between Day 3 and Day 4, while a moderate and significant correlation exists between Day 3 and Day 8.<br>
•	The Nutrient Treatment also demonstrates a strong and significant correlation between Day 3 and Day 4, but the correlation between Day 3 and Day 8 lacks significance.<br>
</div>


These results may imply that treatments using water and nutrients have a noticeable impact on the early growth (i.e., the relationship between lengths on Day 3 and Day 4), while the growth length on Day 8 seems less affected by the length on Day 3. 

VD: well, that's simply because once the plants grow too big, there are many variations among individuals. So, the correlation becomes weaker as plants ae growing and until no significant.
